---
title: "Calculating Sample Size for ICC"
author: "Fran Tomlinson"
date: "02/12/2022"
output:
  html_document: default
  pdf_document: default
---



## Introduction

Using the package "ICC.Sample.Size" [^1], we can determine the required sample size for calculating intraclass correlation coefficient in repeatability studies. 

| ICC         | Interpretation      |
| :---------- | :------------------ |
| <0.5        | Poor agreement      |
| 0.5 to <0.75| Moderate agreement  |
|0.75 to <0.9 | Good agreement      |
|0.9 to 1.0   | Excellent agreement |


```{r - Calculating Sample Size for ICC by Desired Power}
library(ICC.Sample.Size) # Used to calculate sample size for ICC

# calculateIccSampleSize(p,p0,k,alpha,tails,power,by,step)

# p      hypothesised value of ICC
# p0     null hypothesis value of p
# k      number of ratings per subject
# alpha  desired alpha value for hypothesis testing (default 0.05)
# tails  number of tails
# power  desired power of test (default 0.8) 

```


### Inter-rater repeatability

```{r - Inter-Rater Repeatability}
   
# 2 raters per subject
# Hypothesised p=0.9 (excellent ICC) versus null hypothesis p0=0.5 (poor ICC)

calculateIccSampleSize(0.9, 0.5, 2, 0.05,2,0.8,"both")

```


### Intra-Rater Repeatability

```{r - Intra-Rater Repeatability}

# 3 ratings per subject
# Hypothesised p=0.9 (excellent ICC) versus null hypothesis p0=0.5 (poor ICC)

calculateIccSampleSize(0.9, 0.5, 3, 0.05,2,0.8,"both")

```


Key terms:

* Power: the probability that the test correctly rejects the null hypothesis (sensitivity)

* Alpha: the acceptable probability of a false positive (Type I error)


[^1]: Alasdair Rathbone, Saurabh Shaw and Dinesh Kumbhare (2015). ICC.Sample.Size:
  Calculation of Sample Size and Power for ICC. R package version 1.0.
  https://CRAN.R-project.org/package=ICC.Sample.Size

